---
title: "ESM 244 Lab 3 Part 1: Binary Logistic Regression"
author: "Carlos Simms"
date: "1/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)

# Attach packages:

library(tidyverse) # Default.
library(palmerpenguins) # Default dataset.
library(GGally) # To make pairs plots.
library(broom) # To get tidy model outputs in data frames.
library(jtools) # To make some quick graphs of model outputs.
```

**Specify the data for the initial pairs plots.**

```{r}
penguins %>% 
  select(species, 
         bill_length_mm:sex) %>% 
  ggpairs(aes(color = species)) # Gives alot of different relevant relationships and information between the selected variables.
```

**Create a data subset with only Adelie and Chinstrap penguins.**

```{r}
adelie_chinstrap <- penguins %>% 
  filter(species %in% 
           c("Adelie", 
             "Chinstrap")) %>% 
  mutate(species = fct_drop(species)) # To drop any factor levels where there are no observations for that factor level.
```

**Make a graph to visualize the relationship between those two species.**

```{r}
ggplot(data = adelie_chinstrap, 
       aes(x = body_mass_g, 
           y = flipper_length_mm)) + 
  geom_point(aes(color = sex)) + 
  facet_wrap(~species)
```

**Using binary logistic regression, to find the probability between Adelie or Chinstrap penguins, based on several variables via our model.**

**- body mass**
**- flipper length**
**- penguin sex**

```{r}
adelie_chinstrap_blr <- glm(species ~ body_mass_g + 
                              flipper_length_mm + sex, 
                            data = adelie_chinstrap, 
                            family = "binomial") # To indicate that the link is going to be the log odds and that there are only two outcome variables.
```

**Take a look at the outcomes.**

```{r}
adelie_chinstrap_blr

summary(adelie_chinstrap_blr)
```

**Take a look at a graph that could reflect the outcomes viewed.**

```{r}
ggplot(data = adelie_chinstrap, 
       aes(x = species, 
           y = flipper_length_mm)) + 
  geom_jitter(aes(color = sex))
```

**What are the actual probabilities of being a Chinstrap penguin for each of the existing observations in our 'adelie_chinstrap' data frame.**

```{r}
blr_fitted <- adelie_chinstrap_blr %>% 
  broom::augment(type.predict = "response") # 'broom::augment()' by itself, will return the log odds of chinstrap. 'type.predict...' is the prediction calculated to be the probability of each outcome.
```

**Plotting the fitted probabilities.**

```{r}
ggplot(data = blr_fitted, 
       aes(x = flipper_length_mm, 
           y = .fitted)) + 
  geom_point(aes(color = sex, 
                 pch = species)) + 
  geom_smooth(aes(color = sex), 
              se = FALSE) + 
  labs(x = "Flipper Length (mm)", 
       y = "Probability of Outcome Chinstrap")
```

**Plot model based on different variables.**

```{r}
effect_plot(adelie_chinstrap_blr, 
            pred = flipper_length_mm, # To indicate the predictor variable.
            x.label = "Flipper Length (mm)", 
            y.label = "Probability of Chinstrap", 
            interval = TRUE) # To add a confidence interval.

effect_plot(adelie_chinstrap_blr, 
            pred = body_mass_g, 
            x.label = "Body Mass (g)", 
            y.label = "Probability of Chinstrap", 
            interval = TRUE)
```

**What's the probability that a female penguin weighing 3,410 g with a flipper length of 192 mm, is a Chinstrap penguin?**

```{r}
example_1 <- predict(adelie_chinstrap_blr, 
                     data.frame(sex = "female", 
                                body_mass_g = 3410, 
                                flipper_length_mm = 192), # To specify the predictor variable for specific values.
                     type = "response")

example_1
```

**END**
